---
title: "Exam 3"
author: "Marissa Ohran"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1. Load and clean "FacultySalaries_1995.csv" and re-create graph**
   
    
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(janitor)
library(broom)

# Load dataset
sal <- read_csv("FacultySalaries_1995.csv")

# Clean names
sal <- sal %>% clean_names()

# Tidy data
# select state, tier, and ranked salary columns
sal_rank <- sal %>% 
  select(state,tier,ends_with("_salary"))

# pivot_longer, creating a rank column, adjust rank values
sal_rank <-sal_rank %>% 
  pivot_longer(cols = ends_with("_salary"),
               names_to = "rank",
               values_to = "salary") %>% 
  mutate(rank=case_when(rank == "avg_full_prof_salary" ~ "Full",
                        rank == "avg_assoc_prof_salary" ~ "Assoc",
                        TRUE ~ "Assist"))

# filter out tier VIIB
sal_rank <- sal_rank %>% filter(tier != "VIIB")

# plot
# x = rank, y = salary, boxplot, fill = rank, facet wrap by tier
sal_rank %>% 
  ggplot(aes(x=rank, y=salary, fill=rank)) + 
  geom_boxplot() +
  facet_wrap(~tier) +
  theme_minimal() +
  labs(x="Rank", y="Salary",fill="Rank") +
  theme(axis.text.x = element_text(angle = 45))
```
    
    
**2. Build an ANOVA model and display the summary output.**

```{r message=FALSE, warning=FALSE}
# ANOVA model & summary
mod <- sal_rank %>% 
  aov(data = .,formula = salary ~ state + tier + rank)
summary(mod)
```
  
  
  **3. Load "Juniper_Oils.csv" data. Look at it. Then tidy it!**
  
```{r message=FALSE, warning=FALSE}
# load data
juniper <- read_csv("Juniper_Oils.csv")

# look at data
juniper %>% glimpse()

# tidy data
# select columns I care about
juniper_df <- juniper %>% 
  select(c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene","alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene","cuparene","compound 1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound 2","thujopsenal"), YearsSinceBurn)

# pivot_longer, creating ChemID column
juniper_df <- juniper_df %>% 
  pivot_longer(cols = c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene","alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene","cuparene","compound 1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound 2","thujopsenal"),
               names_to = "ChemID",
               values_to = "Concentration")
```
  
  
**4. Make a graph of the following:**

        x = YearsSinceBurn
        y = Concentration
        facet = ChemicalID (use free y-axis scales)
        geom_smooth
        
```{r message=FALSE, warning=FALSE}
# plot
juniper_df %>% 
  ggplot(aes(x=YearsSinceBurn,y=Concentration)) +
  geom_smooth() +
  facet_wrap(~ChemID, scales = "free")
```

        
**5. Use glm to find which chemicals show concentrations that are significantly affected by "Years Since Burn".**
```{r message=FALSE, warning=FALSE}
#glm model
 mod2 <- juniper_df %>% 
 glm(data = ., formula = Concentration ~ YearsSinceBurn * ChemID)

tidy <- tidy(mod2)

tidy %>% 
  filter(p.value <= 0.05) %>% 
  mutate(term=str_remove(term, "ChemID"))
```



glm(Concentration)
